Report

### 2023-07-07 script/ms_report.Rmd
debug report

---
title: "MSGenova (Clean version) Rmakrdown Report"
author: "Gabrièle Adam"
date: "2022-08-16"
output:
    html_document:
        highlight: tango
        number_sections: no
        theme: default
        toc: yes
        toc_depth: 3
        toc_float:
            collapsed: no
            smooth_scroll: yes
params:
  DP.min: DP.min
  AR.min: AR.min
  result_dir: result_dir
  rmd: "report.Rmd"
---

```{r setup, include=FALSE}
workspace<-params$result_dir
knitr::opts_chunk$set(root.dir=workspace, echo = FALSE , warning=FALSE, error=FALSE, message=FALSE)
```

```{r load_library, echo=FALSE, warning=FALSE, error=FALSE}
library(tidyverse)
library(DT)
library(ggplot2)
library(plyr)
library(dplyr)
library(yaml)
vector_calling<-params$vector

DP.min<-5
AR.min<-0.2

sheet<-params$sample
trimmo<- params$trimmo
flagstat<- params$flagstat
haplotype_caller<- params$haplotype_caller

config_advanced_file<-paste0(workspace,"/00_logs/config_advanced")
config_ad<-read_yaml(config_advanced_file)
```

## Trimming parameters

```{r trimmomatic}
trimmo<-paste0(workspace,"/01_sequence_qc/log/trimmomatic.log")
dat<-read.table(text=gsub(" ","\t",readLines(trimmo)),  header=TRUE)
DT::datatable(dat)
```

## Mapping parameters

### Flagstat results

```{r flagstat}
library(tidyverse)
file <- paste0(workspace,"/02_mapping/flagstat/concatenate_flagstat.txt")
if(file.exists(file)){
    flagstat <- read.table(file, sep="\t", header=FALSE) %>% 
      dplyr::mutate(mapped=V3-V4, unmapped=V2-V3, supplementary=V4, 
                    mapped.ratio=round(mapped/(mapped+unmapped)*100,2), 
                    unmapped.ratio=round(unmapped/(mapped+unmapped)*100,2),
                    supplementary.ratio=round(supplementary/(mapped+unmapped)*100,2)) %>% dplyr::select(-V2, -V3, -V4)
    names(flagstat) <- c("sample", "mapped", "unmapped", "supplementary", "mapped.ratio", "unmapped.ratio", "supplementary.ratio")

@@

DT::datatable(flagstat, options = list(rownames = TRUE, pageLength = 10, scrollX = T, dom = 'tip'),class = 'cell-border stripe')
#count
nbr.reads.melt <- flagstat %>% dplyr::select(sample, mapped, unmapped, supplementary) %>% reshape2::melt(id = "sample", value.name = "nbr.reads")
summary.nbr <- nbr.reads.melt %>% group_by(variable) %>% summarise(median=median(nbr.reads), max=max(nbr.reads), min=min(nbr.reads))
nbr.reads.melt <- left_join(nbr.reads.melt, summary.nbr, by = "variable") %>% mutate(x.label=paste0(variable, "\nmedian=", median, "\nmax=", max, "\nmin=", min))
ggplot(data = nbr.reads.melt, aes(y=nbr.reads, x=x.label)) + geom_boxplot()  + xlab("") + ylab("number of reads") + ggtitle("Boxplot representing the number of mapped/supplementary/unmapped reads")+theme_bw()
#ratio
ratio.reads.melt <- flagstat %>% dplyr::select(sample, mapped.ratio, supplementary.ratio) 
names(ratio.reads.melt) <- c("sample", "mapped", "supplementary")
ratio.reads.melt <- ratio.reads.melt %>% reshape2::melt(id = "sample", value.name = "ratio")
summary.ratio <- ratio.reads.melt %>% group_by(variable) %>% summarise( median=round(median(ratio),2), max=round(max(ratio),2), min=round(min(ratio),2))
ratio.reads.melt <- left_join(ratio.reads.melt, summary.ratio, by = "variable") %>% mutate(x.label=paste0(variable, "\nmedian=", median, "%\nmax=", max, "%\nmin=", min, "%"))
ggplot(data = ratio.reads.melt) + geom_boxplot(aes(y=ratio, x=x.label)) + xlab("") + ylab("% reads") + scale_y_continuous(breaks = seq(0, max(ratio.reads.melt$ratio)+10, 20)) + ggtitle("Boxplot representing the percentage of mapped reads and supplementary alignments")+theme_bw()

@@

### Coverage analysis
```{r coverage}
cov.list <- lapply(sample_sheet$sample, function(x){
  file <- paste0(workspace,"/02_mapping/coverage/", x, ".coverage")
  if(file.exists(file)){

@@

cov <- purrr::reduce(cov.list, rbind) %>% dplyr::mutate(nbr_reads=dplyr::if_else(V4 >= 15, ">15X", "<15X"))
colnames(cov) <- c("rname",	"startpos",	"endpos",	"numreads",	"covbases",	"coverage",	"meandepth",	"meanbaseq",	"meanmapq", "sample", "nbr_reads")
#cov <- cov %>% left_join(bed.tab, by=c("rname","startpos"="start","endpos"="stop")) 
cov.summary <- cov %>% group_by(rname) %>% summarise(min=min(numreads), max=max(numreads), median=round(median(numreads)))
cov <- left_join(cov, cov.summary, by = c("rname") )%>% mutate(x.label=paste0(rname, "\nmedian=", median, "\nmax=", max, "\nmin=", min)) 
for (i in unique(sample_sheet$sample)){
cat("####",i,"\n")
g<-ggplot(filter(cov,sample==i)) + geom_boxplot(aes(y=numreads, x=x.label)) + xlab("") + ylab("#reads") + ggtitle("Plot coverage")+facet_wrap(. ~ sample,scales = "free")+theme(axis.text.x=element_text(angle=90,hjust=1))+theme_bw()
print(g)
cat('\n\n')
}
```

### Depth analysis
```{r depth_analysis}
cov.list <- lapply(sample_sheet$sample, function(x){

  file <- paste0(workspace,"/02_mapping/depth/", x, ".depth")
  if(file.exists(file)){
    read.table(file, sep="\t")  %>% dplyr::mutate(plant = x)
  }
})

depth <- purrr::reduce(cov.list, rbind)
colnames(depth) <- c("rname", "position", "depth", "sample")
#depth <- depth %>% left_join(bed.tab, by="rname")  %>% filter(position >=start, position <= stop)
heading<-as.character(unique(sample_sheet$sample))
for (i in unique(cov$rname)){
 cat("####",i,"\n")
  p <- ggplot(data = filter(depth,rname== i)) + geom_line(aes(x=position, y=depth, color=sample))   + theme(legend.position="none", axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + ylab("Depth of coverage")+facet_grid(~rname, scales = "free") 
  print(p)
cat('\n\n')
}

@@

```{r proportion_of, results='asis',echo=FALSE, eval=FALSE, warning=FALSE, error=FALSE,results=FALSE}
for (i in unique(bed.tab$target)){
 cat("####",i,"\n")
tmp <- cov  %>% filter(target== i) %>% group_by(nbr_reads, target, amplicon) %>% count(name = "#well") 
g1<-ggplot(data = tmp) + geom_bar(aes(x=amplicon, y=`#well`, fill=nbr_reads), stat = "identity") + xlab("") +facet_grid(~target, scales = "free")
g2<-cov %>% filter(target== i) %>% group_by(well, nbr_reads,target) %>% count(name = "amplicon") %>% filter(nbr_reads == ">15X") %>% group_by(amplicon,target) %>% count(name = "#well") %>% ggplot(aes(x = amplicon, y = `#well`)) + geom_bar(stat = "identity") + xlab("nbr of amplicon covered at >15X")
print(g1)
print(g2)
cat('\n\n')
}
```

## Study of SNV and Indels calling

```{r define_r_fct_multiallele, echo=FALSE}
vcfR_split_multiAlt <- function(vcfR.tid, nbrAlt=NULL){

  vcfR.tid.tmp <- 
    dplyr::mutate(vcfR.tid, nb.allele=unlist(lapply(str_split(string = vcfR.tid$gt_AD, pattern = ","), function(x){return(length(x))}))) %>%
    dplyr::mutate(nb.allele=nb.allele-1)

  vcfR.tid.list <- lapply(unique(vcfR.tid.tmp$nb.allele), function(x){

    tmp <- vcfR.tid.tmp %>% dplyr::filter(nb.allele == x) %>%
      tidyr::separate(col = gt_AD, into = c("gt_ref_AD", paste0("gt_alt_AD",1:x)), sep = ",", remove = FALSE) %>%
      tidyr::separate(col = ALT,   into = paste0("ALT" ,1:x), sep = ",", remove = FALSE) %>%
      tidyr::separate(col = AC,    into = paste0("AC" ,1:x),  sep = ",", remove = FALSE) %>%
      tidyr::separate(col = AF,    into = paste0("AF" ,1:x),  sep = ",", remove = FALSE) %>%
      tidyr::separate(col = MLEAF, into = paste0("MLEAF" ,1:x),  sep = ",", remove = FALSE) %>%
      tidyr::separate(col = MLEAC, into = paste0("MLEAC" ,1:x),  sep = ",", remove = FALSE)

    tmp.list <- lapply(1:x, FUN = function(i){

        tmp.split <- dplyr::mutate(tmp,  gt_alt_AD = as.numeric(get(paste0("gt_alt_AD", i))) ,
                                         gt_AR     = as.numeric(get(paste0("gt_alt_AD", i)))/gt_DP, 
                                         ID_old    = ID,
                                         gt_AD_old = gt_AD, gt_AD = paste0(gt_ref_AD, ",", get(paste0("gt_alt_AD", i))),
                                         ALT_old   = ALT,   ALT   = get(paste0("ALT", i)),
                                         AC_old    = AC,    AC    = get(paste0("AC", i)),
                                         AF_old    = AF,    AF    = get(paste0("AF", i)),
                                         MLEAF_old = MLEAF, MLEAF = get(paste0("MLEAF", i)),
                                         MLEAC_old = MLEAC, MLEAC = get(paste0("MLEAC", i)),
                                         ID = paste(CHROM, POS, REF, ALT, sep="_"),
                                   ) %>%
        dplyr::select(-gt_ref_AD, 
                      -paste0("ALT", 1:x), 
                      -paste0("gt_alt_AD", 1:x),  
                      -paste0("AC", 1:x),
                      -paste0("AF", 1:x),
                      -paste0("MLEAF", 1:x),
                      -paste0("MLEAC", 1:x)) %>%
        dplyr::filter(ALT != "*")

        })

        return(tmp.list %>% purrr::reduce(rbind))
    })

    return(purrr::reduce(vcfR.tid.list, rbind))
}
```

```{r , echo=FALSE, warning=FALSE, error=FALSE}
haplotype_caller<-paste0(workspace,"/03_snv_indels_calling/",config_ad$project_name,"_snv_indel.vcf")
vcfR.inst <- vcfR::read.vcfR(haplotype_caller)

vcfR.tid.list <- vcfR::vcfR2tidy(vcfR.inst)

vcfR.tid <- full_join(vcfR.tid.list$fix, vcfR.tid.list$gt, by = c("ChromKey", "POS")) %>% 
            dplyr::mutate(ID = dplyr::if_else(is.na(ID), paste(CHROM, POS, REF, ALT, sep="_"), ID))

variants <- vcfR_split_multiAlt(vcfR.tid) %>%
  dplyr::mutate(gt_AR_corr = dplyr::if_else(gt_AR < AR.min, 0, gt_AR),
         gt_AR_corr = dplyr::if_else(gt_DP < 4,   0, gt_AR_corr),
         gt_DP_corr = dplyr::if_else(gt_DP < 4,   0, as.numeric(gt_DP)),
         var_type   = dplyr::if_else(str_length(REF) == str_length(ALT), "SNV", dplyr::if_else(str_length(REF) > str_length(ALT), "DEL", "INS")),
         var_length = str_length(ALT) - str_length(REF),
         var_class  = dplyr::if_else(var_length == 0, var_type, 
                      dplyr::if_else(abs(var_length) == 1,  paste0(var_type, "1"),
                      dplyr::if_else(abs(var_length) >  50, paste0(var_type, ">50"), paste0(var_type, "<50")))))

variants.control <- variants %>% filter(Indiv %in% dplyr::filter(sample_sheet, condition == "WT")$plant, gt_AR_corr > 0) 
variants.freq_corr <- variants %>% filter(gt_AR_corr > 0, !is.na(gt_AR_corr)) %>% 
  select(ID, Indiv, target,amplicon) %>% unique() %>% group_by(ID, target,amplicon) %>% count(name="freq")

variants <- variants %>% dplyr::filter(!ID %in% variants.control$ID) %>% dplyr::full_join(variants.freq_corr, by = c("ID", "target","amplicon"))
#DT::datatable(variants)
```

>>>>

---
title: "Report"
author: " "
date: "2022-08-16"
output:
    html_document:
        highlight: tango
        number_sections: yes
        theme: default
        toc: yes
        toc_depth: 3
        toc_float:
            collapsed: yes
            smooth_scroll: yes
params:
  DP.min: DP.min
  AR.min: AR.min
  qual.min: qual.min
  result_dir: result_dir
  sample_sheet: sample_sheet
  projectName : projectName
  rmd: "report.Rmd"
  AR: AR
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE , warning=FALSE, error=FALSE, message=FALSE, out.width="70%", fig.show="hold")
```

```{r}
# title: "`r params$title`"
workspace <-params$result_dir
sample_sheet.file <- params$sample_sheet
DP.min <- params$DP.min
AR.min <- params$AR.min
qual.min <- params$qual.min
projectName <- params$projectName
#Genome.file <- "/save/project/ijpb/bank/GENOME/Ppatens-patens-v3.3-WG/Ppatens-patens-v3.3-WG.fa"
#projectName <- "PE-WGS"

#workspace <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/results"
#workspace <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/"

#sample_sheet.file <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/sample_sheet_PE.txt"
#sample_sheet.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/sample_sheet_soj_3samples.txt"

# DP.min <- 4
# AR.min <- 0.8
# qual.min <- 30
nbr.alt.min <- 3
```

```{r}
#libraries
library(DT)
library(ggplot2)
library(vcfR)
library(dplyr)
library(tidyr)
library(yaml)
library(stringr)
library(ComplexHeatmap)
library(dendextend) 
library(GenomicRanges)
library(gplots)
source("functions.R")
```

# Objective

# Input data
## Genome version
```{r, eval=FALSE}
print(basename(Genome.file))
```

## Samples
```{r}
# error message if sample sheet not exist
if(!file.exists(sample_sheet.file)) stop("Error: ")

# read sample sheet
sample_sheet <- read.csv(file = sample_sheet.file, header = TRUE, sep = "\t")
```

```{r}
sampleNames <- unique(sample_sheet$sample)
```

```
List of samples : `r sampleNames`
```

```{r list samples}

dplyr::select(sample_sheet, sample, condition) %>% unique() %>% 
  dplyr::group_by(condition) %>% dplyr::count(name = "nb_samples") %>% 
  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
                class = 'cell-border stripe', caption = "List of conditions")

# MsGenoVa workflow

![alt text here](workflow.png)

# Sequence trimming step

Sequences were trimmed with Trimmomatic tools...

```{r trimming stat}
trim.file <- paste0(workspace,"/01_sequence_qc/log/trimmomatic.log")
if(!file.exists(trim.file)) stop("Error: ")

read.table(text = gsub(" ","\t",readLines(trim.file)),  header=TRUE) %>%
  DT::datatable(rownames = FALSE, caption = "Read cleaning results",
                options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
                class = 'cell-border stripe')

# Mapping step

Sequences were aligned using BWA MEM (v?) ...

```{r mapping step}
map.file <- paste0(workspace,"/02_mapping/flagstat/concatenate_flagstat.txt")
if(!file.exists(map.file)) stop("Error: ")
```

## Mapping rate

```{r mapping rate}
flagstat.file <- paste0(workspace,"/02_mapping/flagstat/concatenate_flagstat.txt")
if(file.exists(flagstat.file)){
    flagstat <- read.table(flagstat.file, sep="\t", header=FALSE) %>%
      dplyr::mutate(mapped.reads=V3-V4, mapped.read.ratio=round(mapped.reads/(V2)*100,2)) %>% dplyr::select(-V3, -V4)
    names(flagstat) <- c("sample", "raw.reads", "mapped.reads", "mapped.read.ratio")

@@

#count
nbr.reads.melt <- flagstat %>% dplyr::mutate(xlabel=paste0(sample, "(", mapped.read.ratio, "%)"), unmapped.reads=raw.reads-mapped.reads) %>% 
  dplyr::select(mapped.reads, unmapped.reads, xlabel) %>%
  reshape2::melt(id = "xlabel", value.name = "nbr.reads")

@@

```{r mapping rate plot}
p <- ggplot(data = nbr.reads.melt) + geom_col(aes(y=xlabel, x=nbr.reads, fill=variable)) + 
  labs(x="nb of reads", y="samples (mapping rate %)", fill="")

print(p)
```

```{r mapping rate summary}
DT::datatable(flagstat, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
              class = 'cell-border stripe', rownames = FALSE, caption = "")
```

## Mapping coverage

```{r mapping coverage}
cov.list <- lapply(sample_sheet$sample, function(x){
  file <- paste0(workspace,"/02_mapping/coverage/", x, ".coverage")
  if(file.exists(file)){

@@

cov <- purrr::reduce(cov.list, rbind)
colnames(cov) <- c("rname",	"startpos",	"endpos",	"numreads",	"covbases",	"coverage(%)",	"meandepth(X)",	"meanbaseq",	"meanmapq", "sample")
cov$coverage <- round(cov$coverage, 2)

cov <- cov %>% mutate(xlabel=paste0(rname, "(", endpos-startpos, ")")) %>% 
  dplyr::filter(!stringr::str_detect(stringr::str_to_lower(rname), pattern = "scaffold")) %>%
  dplyr::filter(!stringr::str_detect(stringr::str_to_lower(rname), pattern = "chloroplast")) %>%
  dplyr::filter(!stringr::str_detect(stringr::str_to_lower(rname), pattern = "mitochondria"))
```

```{r mapping coverage plot}
p <- select(cov, rname, `meandepth(X)`, `coverage(%)`, sample, xlabel) %>% reshape2::melt(id = c("rname", "sample", "xlabel")) %>% 
  ggplot() + geom_boxplot(aes(x=value, y=xlabel)) + 
  xlab("") + ylab("chromosome names (size bp)") + 
  facet_grid(~variable, scales = "free")

print(p)
```
```{r mapping coverage summary}
cov %>% dplyr::select(sample, rname, `meandepth(X)`, `coverage(%)`) %>% 
  DT::datatable(rownames = FALSE, caption = "",
                options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
                class = 'cell-border stripe')
```

# Small variant calling step (GATK)
```{r}
vcf.file <- list.files(path = paste0(workspace,"/03_snv_indels_calling"), pattern = "_snv_indel.vcf" , full.names = T)
if(!file.exists(vcf.file)) stop("Error: ")
```

```{r}

vcfR <- vcfR::read.vcfR(vcf.file, verbose = FALSE)

vcfR.tb.list <- vcfR::vcfR2tidy(vcfR, verbose = FALSE, info_fields = c("AC","MQ", "ANN"), 
                                                     format_fields = c("Indiv", "AD", "DP"))

vcfR.tb <- dplyr::full_join(vcfR.tb.list$fix, vcfR.tb.list$gt, by = c("ChromKey", "POS"))
rm(vcfR.tb.list)

# split multiallelic variants + add columns  with var type information
variants.gatk <- vcfR_split_multiAlt(vcfR.tb, nbrAlt = 2)

variants.gatk <- variants.gatk %>%
  tidyr::separate(gt_AD, c("gt_ref_AD", "gt_alt_AD"), sep = ",") %>%
  dplyr::mutate(gt_AR = dplyr::if_else(gt_DP == 0 | is.na(gt_DP), NA, as.numeric(gt_alt_AD)/gt_DP),
                var_length = str_length(ALT) - str_length(REF) + 1,
                var_type   = dplyr::if_else(var_length == 1, "SNV", 
                             dplyr::if_else(var_length  < 1, "DEL", "INS")),
                ID = paste(CHROM, POS, REF, ALT, sep = "_"), tools = "GATK")

variants.gatk <- variant.freq(variants.gatk)

@@

## Variant calling quality control

### Score quality

```{r out.width="60%", fig.show="hold"}
# Score de qualité
p <- variants.gatk %>% select(ID, QUAL) %>% unique() %>% ggplot() + geom_density(aes(QUAL)) + xlim(0,200) +
  xlab("quality score") + geom_vline(xintercept = qual.min, color = "red") + ggtitle("Variant quality score distribution")
print(p)

### Variant coverage

```{r out.width="50%", fig.show="hold"}
vcf.tbl.g   <- dplyr::filter(variants.gatk,   gt_AR != 0, !is.na(gt_AR))

# distribution de la couverture des positions variantes
p <- ggplot(data=vcf.tbl.g) + geom_density(aes(x=gt_DP, color=Indiv)) + xlim(0,50) +
  ylab("DP of variants") + geom_vline(xintercept = DP.min, color = "red") + ggtitle("Variant depth of coverage distribution")
if (length(unique(vcf.tbl.g$Indiv)) > 15){
  p <- p + theme(legend.position="none")
}

print(p)

# Distribution de l'allèle ratio
p <- ggplot(data=vcf.tbl.g) + geom_density(aes(gt_AR, color=Indiv)) + 
  xlab("Alternative allele ratio") + xlim(0,1) + geom_vline(xintercept = AR.min, color = "red")+ ggtitle("Variant allele ratio distribution")

if (length(unique(vcf.tbl.g$Indiv)) > 15){
  p <- p + theme(legend.position="none")
}

print(p)

```{r out.width="60%", fig.show="hold"}
NA_ID <- filter(variants.gatk, nb_NA > 0)$ID

variants.gatk %>% filter(ID %in% NA_ID) %>% 
  ggplot() +  geom_bar(aes(x=as.factor(nb_NA))) + 
  theme(axis.text.x = element_text(angle = 90, size = rel(0.8), hjust = 0)) + xlab("nb of samples") + 
  ylab("Non covered positions") + ggtitle("Non covered variant positions")

### Low quality variant filtering

* score de qualité : > `r qual.min`
* nombre de reads qui couvrent la position variant : > `r DP.min`
* nombre de reads qui portent l'allèle alternatif : > 1
* min AR : `r AR.min`
* retirer les positions avec au moins un échantillon non couvert.

```{r}
# remove 
variants.gatk.flt <- variants.gatk %>% 
  dplyr::mutate(gt_alt_AD = dplyr::if_else(as.numeric(gt_alt_AD) < 2, 0, as.numeric(gt_alt_AD)),
                gt_AR     = dplyr::if_else(gt_DP == 0 | is.na(gt_DP), NA, as.numeric(gt_alt_AD)/gt_DP)) %>% 
  dplyr::filter(!is.na(gt_DP), gt_DP >= DP.min, QUAL >= qual.min, nb_NA == 0, freq_ho == freq)

## Descriptive stat of variants

### variant count per sample

```{r }

vcf.tbl.q.g <- filter(variants.gatk.flt, gt_AR != 0, !is.na(gt_AR))

vcf.tbl.q.g %>% 
  dplyr::mutate(tag=if_else(freq == length(unique(sample_sheet$sample)), "common", "noCommon")) %>% 
  dplyr::select(ID, Indiv, var_type, tag) %>%
  dplyr::mutate(Indiv=if_else(tag == "common", "common", Indiv)) %>% unique() %>%
  ggplot() + geom_bar(aes(x=Indiv, fill=var_type)) + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) + 
  xlab("") + ylab("nb of variants")

```
* common : les variants communs à tous les échantillons.

### Variant frequency in samples

```{r }

p <- vcf.tbl.q.g %>% dplyr::select(Indiv, freq, ID) %>% group_by(Indiv, freq) %>% 
  dplyr::count(name = "nb_var")  %>% filter(freq != length(unique(sample_sheet$sample))) %>% 
  ggplot(aes(x=as.factor(freq), y=nb_var)) + geom_boxplot(outlier.shape = NA) +
  geom_point(size = 0.8, aes(color=Indiv), position=position_jitter(width = 0.2)) + xlab("freq in samples") + ylab("# variants")

if (length(unique(vcf.tbl.q.g$Indiv)) > 15){

  p <- p + theme(legend.position="none")
}
print(p)

# warning -> samples with the same variants or pas de variant
```

### Substitution

```{r}

p <- vcf.tbl.q.g %>% dplyr::filter(var_type == "SNV", str_length(ALT) == 1) %>% 
  dplyr::select(ALT, REF, Indiv) %>% group_by(ALT, REF, Indiv) %>% 
  dplyr::count(name = "nb_sub") %>% 
  ggplot(aes(x=paste(REF, ALT, sep=">"), y=nb_sub)) + geom_boxplot(outlier.shape = NA) +
  geom_point(size = 0.8, aes(color=Indiv), position=position_jitter(width = 0.2)) + 
  facet_grid(~paste0("REF : ",REF), scales = "free") + labs(x="", y="nb of variants", color="samples")

if (length(unique(vcf.tbl.q.g$Indiv)) > 15){

  p <- p + theme(legend.position="none")
}
print(p)

### Clustering per condition

```{r}
vcf.tbl.q.g.condition <- vcf.tbl.q.g %>% 
  dplyr::full_join(dplyr::select(sample_sheet, sample, condition), by = c("Indiv" = "sample")) %>% 
  tidyr::separate(ANN, c("tmp1", "effect", "tmp2", "gene"), sep = "[|]", extra = "drop") %>% dplyr::select(-tmp1, -tmp2)

tb <- vcf.tbl.q.g.condition %>%  
  dplyr::select(ID, condition) %>% 
  dplyr::group_by(ID, condition) %>% unique() %>% 
  reshape2::dcast(ID~condition, value.var = "condition") %>%
  dplyr::mutate_at(.vars = names(.)[-1], .funs = function(x){if_else(is.na(x), 0, 1)})

venn::venn(tb[,-1])
```

### Clustering per sample

```{r }

df <- vcf.tbl.q.g %>% dplyr::filter( freq != length(unique(sample_sheet$sample))) %>% dplyr::select(ID, gt_AR, Indiv) %>% 
  reshape2::dcast(ID~Indiv , value.var = "gt_AR")

# calcul la distance entre les lignes de la matrice
# vcf.tidy.bin[,-1] : sans la 1ere colonne qui contient les IDs
dist  <- dist(t(df[,-1]), method = "euclidean")

# clustering hiérarchique
hc    <- hclust(dist)

# arranger la structure du dendrogramme
dend  <- as.dendrogram(hc)

col.rainbow <- colorspace::rainbow_hcl(length(unique(sample_sheet$condition)))

col <- col.rainbow[as.factor(sample_sheet$condition)]
names(col) <- sample_sheet$sample

labels_colors(dend) <- col[order.dendrogram(dend)]
# dessiner le dendrogramme
plot(dend, main = "Sample clustering using all variants")
```

```{r, eval=FALSE}
## annotations

colname.ANN <- c("Allele", "effect", "impact", "GeneName", 
                 "GeneID", "FeatureType", "FeatureID", "TranscriptBiotype", 
                 "Rank", "HGVS.c", "HGVS.p", "cDNA_len:", 
                 "CDS_len", "Protein_len", "DistanceToFeature", "Warnings")

exlude_effect <- c("upstream_gene_variant", "downstream_gene_variant", "intergenic_region", "intragenic_variant", "intragenic_variant", 
                   "3_prime_UTR_variant", "", "", "")

ID.ANN <- vcf.tbl.q.g %>% dplyr::select(ID, ANN) %>% unique()

ANN.split <- lapply(1:dim(ID.ANN)[1], FUN = function(x){

  tmp <- ID.ANN$ANN[x] %>% str_split(",") %>% unlist()
  data.frame(ID = rep(ID.ANN$ID[x], length(tmp)), ANN = tmp)

}) %>% purrr::reduce(rbind) %>% 
  tidyr::separate(col = ANN, into = colname.ANN, sep = "[|]") %>% 
  dplyr::filter(!effect %in% exlude_effect) %>% 
  dplyr::select(ID, effect) %>% unique()

effect.split <- lapply(1:dim(ANN.split)[1], function(x){

  tmp <- ANN.split$effect[x] %>% stringr::str_split("&") %>% unlist()
  data.frame(ID = rep(ANN.split$ID[x], length(tmp)), effect = tmp)
}) %>% purrr::reduce(rbind)

effect.split %>% dplyr::group_by(effect) %>% dplyr::count()

### 2023-07-07 script/report.Rmd
debug report

---
title: "Report"
author: " "
date: "2022-08-16"
output:
    html_document:
        number_sections: yes
params:
  DP.min: DP.min
  AR.min: AR.min
  qual.min: qual.min
  result_dir: result_dir
  sample_sheet: sample_sheet
  rmd: "report.Rmd"
  AR: AR
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE , warning=FALSE, error=FALSE, message=FALSE)
```

```{r}
# title: "`r params$title`"
# workspace<-params$result_dir
# sample_sheet.file <- params$sample_sheet
# DP.min <- params$DP.min
# AR.min <- params$AR.min
# qual.min <- params$qual.min

projectName <- "PE-WGS"

workspace <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/results"
#workspace <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/"

sample_sheet.file <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/sample_sheet_PE.txt"
#sample_sheet.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/sample_sheet_soj_3samples.txt"

DP.min <- 4
AR.min <- 0.8
qual.min <- 30

# results
trim.file <- paste0(workspace,"/01_sequence_qc/log/trimmomatic.log")
vcf.file <- paste0(workspace,"/03_snv_indels_calling/", projectName , "_snv_indel.vcf")
sv.file  <- paste0(workspace,"/04_sv_calling/", projectName , "_sv.vcf")

if(!file.exists(sample_sheet.file)) stop("Error: ")
if(!file.exists(trim.file)) stop("Error: ")
if(!file.exists(vcf.file)) stop("Error: ")

eval.SV <- file.exists(sv.file)

@@

```{r }
library(DT)
library(ggplot2)
library(vcfR)
library(dplyr)
library(tidyr)
library(yaml)
library(stringr)
library(ComplexHeatmap)
library(gplots)
source("functions.R")

# Samples
```{r}

sample_sheet <- read.table(file = sample_sheet.file, header = TRUE, sep = "\t")

unique(sample_sheet$sample)
```
List of samples : `r unique(sample_sheet$sample)`

```{r}

dplyr::select(sample_sheet, sample, condition) %>% unique() %>% dplyr::group_by(condition) %>% dplyr::count(name = "nb_samples") %>% 
  knitr::kable(align = "lccrr")
 # DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 10, scrollX = T, dom = 'tip'),class = 'cell-border stripe')

<!-- # Quality control -->

```{r quality_control}

  out <- NULL

  out <- c(out, knitr::knit_child("quality_control.Rmd"), quiet=TRUE)

  knitr::asis_output(out)

<!-- # SNV and Small InDels Detection -->

```{r SNV_InDels_calling}

  out <- NULL

  out <- c(out, knitr::knit_child("SNV_InDels_calling.Rmd"), quiet=TRUE)

knitr::asis_output(out)

<!-- # SV Detection -->

```{r SV_calling, eval=eval.SV}

  out <- NULL

  out <- c(out, knitr::knit_child("SV_calling.Rmd"), quiet=TRUE)

knitr::asis_output(out)

```{r tDNA_scan, eval=FALSE, include=FALSE}

  out <- NULL

  out <- c(out, knitr::knit_child("tDNA_scan.Rmd"), quiet=TRUE)

knitr::asis_output(out)

```{r tools, eval=FALSE}

  out <- NULL

  out <- c(out, knitr::knit_child("tools.Rmd"), quiet=TRUE)

knitr::asis_output(out)

>>>>

---
title: "Report"
author: " "
date: "2022-08-16"
output:
    html_document:
        highlight: tango
        number_sections: yes
        theme: default
        toc: yes
        toc_depth: 3
        toc_float:
            collapsed: yes
            smooth_scroll: yes
params:
  DP.min: DP.min
  AR.min: AR.min
  qual.min: qual.min
  result_dir: result_dir
  sample_sheet: sample_sheet
  projectName : projectName
  rmd: "report.Rmd"
  AR: AR
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE , warning=FALSE, error=FALSE, message=FALSE, out.width="70%", fig.show="hold")
```

```{r}
# title: "`r params$title`"
workspace <-params$result_dir
sample_sheet.file <- params$sample_sheet
DP.min <- params$DP.min
AR.min <- params$AR.min
qual.min <- params$qual.min
projectName <- params$projectName
#Genome.file <- "/save/project/ijpb/bank/GENOME/Ppatens-patens-v3.3-WG/Ppatens-patens-v3.3-WG.fa"
#projectName <- "PE-WGS"

#workspace <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/results"
#workspace <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/results/"

#sample_sheet.file <- "/Users/nbentahar/Desktop/msgenova_test/PE_WGS/sample_sheet_PE.txt"
#sample_sheet.file <- "/Users/nbentahar/Documents/PROJECTs/DEV/msgenova/examples/soj/sample_sheet_soj_3samples.txt"

# DP.min <- 4
# AR.min <- 0.8
# qual.min <- 30
nbr.alt.min <- 3
```

```{r}
#libraries
library(DT)
library(ggplot2)
library(vcfR)
library(dplyr)
library(tidyr)
library(yaml)
library(stringr)
library(ComplexHeatmap)
library(dendextend) 
library(GenomicRanges)
library(gplots)
source("functions.R")
```

# Objective

# Input data
## Genome version
```{r, eval=FALSE}
print(basename(Genome.file))
```

## Samples
```{r}
# error message if sample sheet not exist
if(!file.exists(sample_sheet.file)) stop("Error: ")

# read sample sheet
sample_sheet <- read.csv(file = sample_sheet.file, header = TRUE, sep = "\t")
```

```{r}
sampleNames <- unique(sample_sheet$sample)
```

```
List of samples : `r sampleNames`
```

```{r list samples}

dplyr::select(sample_sheet, sample, condition) %>% unique() %>% 
  dplyr::group_by(condition) %>% dplyr::count(name = "nb_samples") %>% 
  DT::datatable(rownames = FALSE, options = list(rownames = TRUE, pageLength = 6, scrollX = T, dom = 'tip'),
                class = 'cell-border stripe', caption = "List of conditions")

@@

# MsGenoVa workflow

![alt text here](workflow.png)
